---
title: "Introduction to Grapher"
output: html_notebook
---

```{r setup, include = FALSE}
library("igraph")
library("grapher")
```
## Overview

The `grapher` package provides and extends an R-interface to the open-source [grapher library]() by Ayasdi. 
It supports interactive [WebGL](https://www.khronos.org/webgl/)-accelerated 2D network plotting and interaction via the [htmlwidgets framework](https://github.com/ramnathv/htmlwidgets). 

## Usage

To create a grapher instance, pass in one of the following: 
  
  1. An [igraph]() object 
  2. An $n \times n$ symmetric adjacency matrix
  3. An $n \times 2$ edgelist 
  4. A [grapher configuration](?grapher::getDefaultJsonConfig)
  
Here's an example of how to create simple graph with 10 nodes. 
```{r}
require("igraph")
n_nodes <- 10L
G <- igraph::sample_grg(n_nodes, radius = 0)
grapher(G)
```

The grapher instance itself encapsulates canvas-drawn htmlwidget that can be programmatically interacted with. You can left-click anywhere in the space to pan, or click an individual node to drag it. Zooming is also available via the [wheel input](https://developer.mozilla.org/en-US/docs/Web/API/GlobalEventHandlers/onwheel). 

One such way of interacting with the widget is through method chaining via [magrittr pipes](https://magrittr.tidyverse.org/). For example, to update the node colors, the `updateNodeColor` accepts a vector of hexadecimal RGB or RGBa values with the same length as the number of nodes:
```{r}
{ light_red <- rgb(1, 0, 0, 0.5); light_green <- rgb(0, 1, 0, 0.5) }
node_colors <- rep(c(light_red, light_green), times = n_nodes/2)
grapher(G) %>% 
  setNodeColor(color = node_colors)
```

Edge colors are updated synonomously. Unlike nodes, edge colors do not support transparancy values. 
```{r}
blue <- rgb(0, 0, 1); red <- rgb(1, 0, 0)
node_colors <- rep(red, igraph::vcount(G))
edge_colors <- rep(blue, igraph::ecount(G))
grapher(G) %>% 
  setNodeColor(color = node_colors) %>% 
  setLinkColor(color = edge_colors)
```

By default, `grapher` instances have their layouts precomputed using igraph before being displayed in their 'static' positions. Alternatively, one can enable a [force-directed](https://en.wikipedia.org/wiki/Force-directed_graph_drawing) layout using [D3v4](https://d3js.org/) with some suitable default parameters via the `enableForce` option.    

```{r}
grapher(G) %>% enableForce() %>% center()
```

The package binds API calls directly with the widget instance, and provides reactive callbacks to work with [shiny](https://shiny.rstudio.com/), enabling more dynamic interactivity with the widget. 

```{r}
library(shiny)
ui <- pageWithSidebar(
  headerPanel("Simple shiny app with Grapher"),
  sidebarPanel(
    actionButton(inputId = "add_node", label = "Add node"), 
    actionButton(inputId = "remove_node", label = "Remove node")
  ),
  ## The main canvas container holding the grapher
  mainPanel( grapherOutput(outputId = "grapher", height = "100vh") )
)

server <- function(input, output, session) {
  n <- igraph::vcount(G)
  
  ## Pass the htmlwidget to the canavs container via 'renderGrapher'
  output$grapher <- renderGrapher({ grapher(G) %>% enableForce() })
  
  ## See ?getDefaultJsonConfig or ?insertNodes to understand the format grapher expects when inserting nodes
  observeEvent(input$add_node, {
    node_config <- data.frame(id=n, x=runif(1), y=runif(1), r=5, color="#000000")
    n <<- n + 1
    insertNodes("grapher", node_config)
  })
  observeEvent(input$remove_node, {
    n <<- n - 1
    removeNodes("grapher", n)
  })
}

shinyApp(ui, server)
```

```{r, include = FALSE}
# label_xy <- data.frame(label = 1:2, x = c(0.05, 0.95), y = c(0.5, 0.5))
# node_labels <- c(rep(1, n_nodes/2), rep(2, n_nodes/2))
# 
# grapher(G) %>% 
#   updateNodeColor(color = c(rep(rgb(1, 0, 0, 0.5), n_nodes/2), rep(rgb(0, 1, 0, 0.5), n_nodes/2))) %>% 
#   splitByLabel(node_labels = node_labels, label_xy = label_xy) %>% 
#   groupSingletons()
# 
# edge_colors <- apply(igraph::as_edgelist(G), 1, function(edge){ 
#   if (edge[1] <= 25 && edge[2] <= 25){ rgb(1, 0, 0) } 
#   else if (edge[1] > 25 && edge[2] > 25){ rgb(0, 1, 0) } 
#   else { rgb(0, 0, 1) }
# })
# grapher(adjm) %>% 
#   updateNodeColor(color = c(rep(rgb(1, 0, 0), n_nodes/2), rep(rgb(0, 1, 0), n_nodes/2))) %>% 
#   updateEdgeColor(color = edge_colors) %>% 
#   grapher:::logNetwork()
# 
# 
# el <- igraph::as_edgelist(G) 
# edge_types <- cbind(node_labels[el[, 1]], node_labels[el[, 2]])
# internal_edges <- which(apply(edge_types, 1, function(e) e[1] == e[2]))
# cross_edges <- which(apply(edge_types, 1, function(e) e[1] != e[2]))
# 
# grapher(G) %>% 
#     updateNodeColor(color = c(rep(rgb(1, 0, 0, 0.5), n_nodes/2), rep(rgb(0, 1, 0, 0.5), n_nodes/2))) %>% 
#     splitByLabel(node_labels = c(rep(1, n_nodes/2), rep(2, n_nodes/2)), label_xy = label_xy) %>% 
#     removeEdges(v1 = el[cross_edges, 1], v2 = el[cross_edges, 2]) %>% 
#     removeEdges(v1 = el[internal_edges, 1], v2 = el[internal_edges, 2]) %>% 
#     addEdges(v1 = el[1, 1], v2 = el[1, 2], color = rgb(1, 0, 0))
# 
# 
# grapher(G) %>% 
#   updateNodeColor(color = c(rep(rgb(1, 0, 0, 0.5), n_nodes/2), rep(rgb(0, 1, 0, 0.5), n_nodes/2))) %>% 
#   splitByLabel(node_labels = c(rep(1, n_nodes/2), rep(2, n_nodes/2)), label_xy = label_xy) %>% 
#   addEdges(v1 = el[cross_edges, 1], v2 = el[cross_edges, 2], color = rgb(0, 0, 1))
#   removeEdges(v1 = el[cross_edges, 1], v2 = el[cross_edges, 2]) %>% 
#   removeEdges(v1 = el[internal_edges, 1], v2 = el[internal_edges, 2]) %>% 
#   addEdges(v1 = el[cross_edges, 1], v2 = el[cross_edges, 2], color = rgb(0, 0, 1)) %>% 
#   addEdges(v1 = el[1:5, 1], v2 = el[1:5, 2], color = rgb(0, 0, 1)) %>% 
#   addEdges(v1 = el[6:10, 1], v2 = el[6:10, 2], color = rgb(0, 1, 0))
#   grapher:::logNetwork()
# 
# 
# net_config <-  grapher::getDefaultJsonConfig(G)
# new_config <- net_config
# new_config$nodes <- new_config$nodes[1,]
# new_config$nodes$color <- rgb(0, 0, 1)
# new_config$links <- data.frame(from = 1, to = 2, color = rgb(0, 0, 1))
# grapher(net_config) %>% 
#   grapher::updateNodes(node_ids = c(1L), net_config = new_config) %>% 
#   grapher:::logNetwork()
#   updateNodeColor(color = c(rep(rgb(1, 0, 0, 0.5), n_nodes/2), rep(rgb(0, 1, 0, 0.5), n_nodes/2))) %>% 
#   splitByLabel(node_labels = node_labels, label_xy = label_xy) %>% 
#   groupSingletons()
```


